---
title: Useful R Packages for Causal Inference in the Social Sciences
author: F. Daniel Hidalgo
date: '2019-09-27'
slug: useful-r-packages-for-causal-inference-in-the-social-sciences
categories:
  - r
  - statistics
  - quant2
tags:
  - r
  - statistics
subtitle: ''
summary: ''
authors: []
lastmod: '2020-04-30'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
draft: false
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(gt)
```

The table below lists `R` packages that I have found useful for causal inference in the social sciences, particularly political science.  I regularly teach [Causal Inference](https://www.dropbox.com/s/80xrhc0shr2npr9/17_802_syll2020.pdf?dl=0)  as part of the methods sequence at MIT Political Science, so this post mainly serves a convenient place to send students when teaching that course. Nevertheless, it may be useful to other applied researchers. 

Of course, the table below is by no means exhaustive, but merely reflects what I've used in the course of research and teaching. Please contact me if you have any additional suggestions or comments. 

```{r, echo=FALSE}
packages <- readxl::read_xlsx("./useful_r_packages.xlsx") %>%
  mutate(Package = glue::glue("[`{Package}`]({Url})"),
          Url = glue::glue("<{Url}>")) %>%
  select(-Url)

gt(packages) %>%
  fmt_markdown(everything()) %>%
  cols_label(
    Package = md("**Package**"),
  #  Url = md("**URL**"),
    Notes = md("**Notes**")
  ) %>%
  tab_options(
    table.font.size = "smaller",
    column_labels.font.size = "small"
  )
```

